Staged Abstract Interpreters

### 2018-11-17 paper/60_eval.tex
evaluation

Our evaluations have shown that the staged abstract interpreter performs well in some
interesting cases that are considered to be the worst case scenario in the control flow
analysis, the staged version outperforms the unstaged abstract interpreter by a wide
margin (up to 3x performance gain).

\subsection{Benchmarks}

Our evaluation is based on a suite of benchmark tests. Considering that our abstract 
interpreters are implemented in Scala which will be affected by JVM warm up times, programs
tend to run faster after a number of runs. In order to minimize the influence of JVM, 
we ran all experiments 10 times and take the statistical mean and median values of 
the running times.

%on a suite of benchmark scheme programs that are also used in \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical}. 
We used the following benchmark programs:
\begin{itemize}
    \item \textbf{fib:} a program that recursively calculates the $n$-th fibonacci number in
        exponential time.
    \item \textbf{church:} a program introduced by \todo{cite Dimitrios Vardoulakis and Olin Shivers. CFA2: a Context-Free approach
        to Control-Flow analysis. Logical Methods in Computer Science,
        7(2), 2011.} to test distributivity of multiplication for numbers in Church Encoding \todo{ cite }
    \item \textbf{kcfa3/kcfa-worst-case-n:} Benchmark programs that are supposed to
      be tough cases for CFA.
    \item \textbf{fermat:} Fermat and Solovay-Strassen primality testing in Scheme by
      Matthew Might.
    \item \textbf{rsa:} The RSA public key encryption algorithm in Scheme.
\end{itemize}

@@

For each benchmark program, we ran our prototype abstract interpreters, both unstaged and staged.
\todo{Figure show data.}

@@

Figure~\ref{evaluation_result} shows the benchmark evaluation results. The "unstaged" and "staged" columns
show the median time to finish the abstract interpretation (in milliseconds). The "$\Delta_{median}$"
column shows the performance change from the unstaged abstract interpreter to the staged version.\\

However, there also exists some programs where the unstaged abstract interpreter yields 
results faster than the unstaged interpreter.
From the data shown above, we observed that 
However, in general, the unstaged abstract interpreter yields results faster
than the unstaged interpreter.

>>>>

Our evaluations shows that the staged abstract interpreter performs well in some
cases that are considered to be the worst scenarios for traditional $k$-CFA.
The staged version outperforms the unstaged abstract interpreter by a significant
margin (up to ~230\% performance gain) on these difficult ones.

\subsection{Benchmarks}
The benchmark programs we used are collected from previous papers \cite{Johnson:2013:OAA:2500365.2500604,
ashley:practical, DBLP:journals/corr/abs-1102-3676} and exisiting artifacts \footnote{https://github.com/ilyasergey/reachability}.

We used the following benchmark programs:
\begin{itemize}
    \item \textbf{fib}: Calculates the $n$-th fibonacci number.
    \item \textbf{rsa}: The RSA public key encryption algorithm.
    \item \textbf{kcfa3}: A difficult benchmark for $k$-CFA.
    \item \textbf{church}: Church numerals with additions and mutiplications.
    \item \textbf{fermat}: Fermat and Solovay-Strassen primality testing.
    \item \textbf{kcfa-worst-case-}\textit{n}: 
      Benchmark programs that are supposed to be tough cases for $k$-CFA; 
      the number $n$ indicates the depth of nesting lambda terms.
\end{itemize}

@@

Considering that our abstract interpreters are implemented in Scala which will be 
affected by the JIT warm up times, we run all experiments 10 times and reprot
the statistical median values of the running times.

@@

Figure~\ref{evaluation_result} shows the benchmark evaluation results. The "unstaged" and "staged" columns
show the median time to finish the analysis (in milliseconds). The "$\Delta$ of median"
column shows the performance change from the unstaged abstract interpreter to the staged version.

As we can see from the table, the staged version significantly outperforms the unstaged especially 
for the difficult ones.
However, for some benchmarks, we observe degraded performance. 
We conjecture this happens due to the large size of the generated code is unoptimized -- 
as what we test here is a simple implementation almost identical to what we described in the previous sections,
we believe the performance still can be improved if more engineering effort is taken, 
such as optimizing code generation and avoiding code duplication.
